---
title: "Scenario Evaluation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation & Cleaning
```{r library, message=F}
library(magrittr)
library(dplyr)
```

## Calculating the Probability of Default for Test Set

## Importing the Test Datasets
```{r}
test_scenario3 <- read_csv("test_scenario3.csv")
test_scenario1_2 <- read_csv("test_scenario1_2.csv")
```

## Getting Columns Specific to Scenario 1, 2 and 3

```{r}
requested_loan_sc1_2 <- test_scenario1_2 %>% dplyr::select(requested_loan)
requested_loan_sc3 <- test_scenario3 %>% dplyr::select(requested_loan)
proposed_intrest_rate <- test_scenario3 %>% dplyr::select(Proposed_Intrest_Rate)
```


## Loading PD and LGD of Customers in Test Set

```{r}
LGD <- read.csv("LGD.csv")
PD <- read.csv("PD.csv")
```


## Evaluating Customers in Scenario 1

CAPITAL <- 1.4 Billion

```{r}
scenario_1 <- test_scenario1_2 %>%
  dplyr::select(id) %>%
  bind_cols(requested_loan_sc1_2, PD, LGD) %>%
  rename(LGD = x) %>%
  mutate(
    Gain = requested_loan * 0.0432 * 5 * (1 - PD),
    Risk = requested_loan * PD *  LGD,
    oppurtunity = Gain- Risk
  )

summary(scenario_1)
```

## Total Requested Loan in Scenario 1
```{r}
scenario_1 %>%
  dplyr::select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```



## PD at 0.5 in Scenario 1
```{r}
scenario_1 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  dplyr::select(default) %>%
  summary()
```

## Total Sum of Gain if PD >=0.5
```{r}
scenario_1 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  dplyr::select(Gain) %>%
  transmute(total_gain = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## PD at 0.1 in Scenario 1
```{r}
scenario_1 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  dplyr::select(default) %>%
  summary()
```

## Total Sum of Gain if PD >=0.1
```{r}
scenario_1 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  dplyr::select(Gain) %>%
  transmute(total_gain = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## Approving Customers in Scenario 1 based on Gain >= Risk
```{r}
scenario_1 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  dplyr::select(loan_approval) %>%
  summary()
```

## Total Sum of Gain in Scenario 1
```{r}
scenario_1 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  dplyr::select(Gain) %>%
  transmute(total_gain = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## Total Requested Loan where Gain >= Risk in Scenario 1
```{r}
scenario_1 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  dplyr::select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## Final Decision on Customer Loan Approval in Scenario 1
```{r}
scenario_1 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  dplyr::select(id,loan_approval) %>%
  write.csv(file = "G6_S1.csv", row.names = F)
```

### Evaluating Customers in Scenario 2

Available loans = 450million
```{r}
scenario_2 <- test_scenario1_2 %>%
  dplyr::select(id) %>%
  bind_cols(requested_loan_sc1_2, PD, LGD) %>%
  rename(LGD = x) %>%
  mutate(
    Gain = requested_loan * 0.0432 * 5 * (1 - PD),
    Risk = requested_loan * PD *  LGD,
    Opportunity = Gain - Risk
  )



```

### Final Decision on Customer Loan Approval in Scenario 2

The second scenario is similar to the first but now we have a budget of $450 million. 

```{r}
Scen_2<-  scenario_2 %>%
  group_by(id, requested_loan) %>%
  summarise(loan_approval = ifelse(Opportunity > 13820, 1, 0))

total_loan <- Scen_2  %>% dplyr::filter(loan_approval == 1)

scen_2_results <- total_loan %>% dplyr::select(id, loan_approval)

```

```{r}
write.csv(file = "G6_S2.csv", row.names = F)
```


```{r}
sum(total_loan$requested_loan) 
```
 

### Evaluating Customers in Scenario 3

Loans available = 1.4 Billion

```{r}
scenario_3 <- test_scenario3 %>%
  dplyr::select(id) %>%
  bind_cols(requested_loan_sc1_2, proposed_intrest_rate, PD, LGD) %>%
  rename(LGD = x) %>%
  mutate(
    Gain = requested_loan * Proposed_Intrest_Rate * 5 * (1 - PD),
    Risk = requested_loan * PD *  LGD
  )

head(scenario_3)
```

```{r}
scenario_3 %>% dplyr::select(Proposed_Intrest_Rate) %>% summary()
```

## Total Requested Loan where Gain >= Risk in Scenario 3

```{r}
scenario_3 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  dplyr::select(loan_approval) %>%
  summary()
```


```{r}
scenario_3 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  dplyr::select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## Total Sum of Gain if Gain >= Risk
```{r}
scenario_3 %>%
  mutate(loan_approval = ifelse(Gain >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  dplyr::select(Gain) %>%
  transmute(total_gain = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## PD at 0.5 in Scenario 3
```{r}
scenario_3 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  dplyr::select(default) %>%
  summary()
```

## Total Sum of Gain if PD >=0.5
```{r}
scenario_3 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  dplyr::select(Gain) %>%
  transmute(total_gain = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

## PD at 0.1 in Scenario 3
```{r}
scenario_3 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  dplyr::select(default) %>%
  summary()
```

## Total Sum of Gain if PD >=0.1
```{r}
scenario_3 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  dplyr::select(Gain) %>%
  transmute(total_gain = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```



### Final Descision on Customer Loan Approval in Scenario 3
```{r}
scenario_3 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate(loan_approval = ifelse(default == 0, 1, 0)) %>%
  dplyr::select(id,loan_approval) %>%
  write.csv(file = "G6_S3.csv", row.names = F)
```

```{r}
scenario_3 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  dplyr::select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```
